Regression Cube Analysis of Cohort Study Data

### 2015-03-31 draft.tex
added revision based on Bernhards notes

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item The analysis of confounding features by providing comparison views between different formula results.
%	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
	\item The expansion and adaption of our method to any data with our open web-based system.
\end{itemize}
%%
The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features using statistical methods.
%%
As described by Thew et al. \cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{itemize}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{itemize}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow. Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data.
%%
The presented work is based on regression analysis as well.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value is in the range [0,1], where 1 encodes a perfect fit.

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
\begin{equation}
Dependent \sim Independent_1 + ... + Independent_n
%Dependent \sim \sum_{i=1}^n Independent_i
\label{eq:RegressionNotation}
\end{equation}
%%
An example of a regression formula would be $KidneyDisorder \sim Smoking + Obesity$.
The most commonly used regression operators comprise:
%%
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable (e.g. $x\pm y$),
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
The type of the target feature restricts the regression type.
%%
Different regression types are available, we focus on the following:
%%
\paragraph{Linear Regression for Continuous Target.} The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
%%
The \emph{dependent} variable has to be of a continuous type.
%%

@@

While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
\begin{enumerate}
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
%%
The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
Overview visualizations of feature relationships as presented by Angelelli et al. \cite{Angelelli} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
These target features often indicate the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing these relationships.
%%
% unified -> singular [Steffi?]
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
\subsection{Cube Description Using Regression Formula Notation}
%%
Expert knowledge is integrated into regression analysis by the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
%%

@@

\subsection{Abstracting Regression Results Using $R^2$}
\begin{figure}[htb]
 \centering
 % \includegraphics[width=3.5in]{figures/cube}
 \includegraphics[width=1.0\linewidth]{figures/cube}
 \caption{
 (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The $R^2$ values extracted from the regression formulas depict the quality-of-fit and are mapped to color saturation (a saturated color shows a strong correlation).
 %%
 (right) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The $R^2$ value of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between [0,1], it allows for comparison between regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%

\paragraph{3D (Cube) View.}
%%
Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube} b).
%%
$R^2$ values of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
The visualization of $R^2$ values derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
% e.g. wenn man es wâ°hlen kann
% i.e. wenn es so ist
Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., blue for linear regression and red for logistic regression).
%%
% Adverb immer vors Hauptverb oder Satzende
Thus, the cube can be easily extended using other regression types.
%%
For cubes with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
The color for both 2D and 3D view encodes the type of the regression analysis (blue for logistic regressions, orange for linear regressions).
%%
\\\\
%%
Our goal is to create an overview visualization for a data set, but we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{Regression Cube}.

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{Regression Cubes} (based on~\cite{Keim}).
 %%
 % The analysis starts with a set of features given by the data set.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare regression cubes.
 %%
 [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
%\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
%\com{Workflow for different analysis approaches.}
%\com{0. Select the Variables.}
%\com{1. Select the Formulas}
%\com{a) Analysis without hypothesis about the data set (z ~ x+y)}
%\com{b) Analysis without hypothesis with confounder (z ~ age)}
%\com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
%\com{d) Use Difference View}
%\com{d) Clarify using confounder (Var ~ x)}
%\com{3. Find interesting pane.}
%\com{4. Get Details on Demand.}
%%
% â¹bergang ist recht holprig gewesen
Our \emph{Regression Cube} model is well-suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al. \cite{Keim}:

@@

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
As seen in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new regression cube for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
Input formulas reflect \emph{hypotheses} about the data.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

%%
$Cancer \sim X + Y + Z$ is an example for the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
%%
All combinations of three independent features towards the target are analyzed through the cube.

@@

\section{System Design} \label{sec:SystemDesign}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/system}
 \caption{
 %%
 Breast fat data set loaded into the \emph{Regression Cube} prototype.
 %%
 (a) Using the formula input, the user specifies the dependent feature as well as its calculation rules.
 %%
 (b) 3D cube visualization, showing values above the cube matrix diagonal as overview.
 %%
 The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
 %%
 (c) 2D mosaic plot visualization of the selected slice for feature \emph{Pain/Discomfort}.
 %%††
 }
  \label{fig:System}
\end{figure}
%% scientific english: kurz und langweilig ;)
We designed our system to be freely accessible and easy to use.
With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
%%
The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.
%%
Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
%%
\\\\
%%
\noindent Design choices and spaces are restricted by the underlying technologies and devices.
%%
Resting method and data exchange on web-based technologies offers many advantages.
%%
There is no set-up time involved and domain experts can use the methods from any computer connected with the web.
%%
By design, web technology is based on a client-server architecture, making it easy to outsource computationally heavy tasks on server clusters and transferring results to the client device.
%%
The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
%%
Therefore, we have to adapt our design respecting the affordances of web pages.
%%
\subsection{System Paradigm and Components}
%%
The \emph{Regression Cube} design focuses on a clean interface, reducing the amount of user-interface elements as much as possible.
%%
This allows for a fast learning of the individual system parts.
%%
%More importantly, it allows the user to focus all mental resources on the analysis tasks, rather than wrangling with configuring the system to serve the current task or hypothesis.
%%
Our prototype consists of three components:
\begin{itemize}
	%%
	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
	%%
	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube [2.SI].
	%%
	\item The \emph{formula editor} allows to adjust the formula w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
	%%
	It also allows to select a reference formula for creating a difference cube [3.ZF].
	%%
\end{itemize}
%%\com{Focus on simplification and ease of use, reducing the number of UI elements as much as possible (Statistical Visualizations often tend to put all information and settings at once to the user, which is overwhelming for the user).}
\paragraph{File Upload and Classification [1.AF].}
%%
Popular analytics tools, such as WEKA \cite{WEKA}, owe their popularity to their support of open file types.
%%
To allow other users even outside the epidemiological application domain to access to our tool, we use standard ASCII-based CSV files.
%%
The first line in a CSV file represents all features (columns) of the data set.
%%
Each line after that represents one subject (row) and its feature manifestations.

\textbf{Encoding via CSV files.}
%%
Encoding variable types in CSV files is not standardized.
%%
However, we need to ensure the correct variable type classification and have to enforce some basic standards.
%%
All categorical values have to be enclosed by quotation marks.
%%
Continuous variables are denoted as digits without enclosing quotation marks.
%%
Although this seems obvious, many cohort study data sets encode categorical features using ID values that are denoted in a data dictionary.
%%
Variables with only two manifestations are classified as dichotomous, leading to three possible data types: numerical, categorical and categorical/dichotomous.
%%
Missing values are denoted by using no character at all, a whitespace, or an empty quotation mark encapsulated string.

@@

\paragraph{Formula Editor [1.AF, 3.ZF].}
%%
After uploading the data, the user can specify a formula or use the default ($Z \sim X + Y$).
%%
Entering a formula is provided via text input.
%%
On formula input, a context modal displays all data set features as well as the available operators and their function.
%%
This allows to comprehend the function of the underlying formula for users without statistical background about regression analysis and its notation.
%%
Auto-completing input features also simplifies the approach and works as spell check of feature names.
%%

@@

\paragraph{Usage of a Regression Prism for information reduction.}
%%
Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the mosaic plot matrix.
%%
This is due to the symmetry of basic regression operators.
%%
$Z \sim X + Y$ produces the same result as $Z \sim Y + X$.
%%
Therefore, we can discard half of the results to reduce visual clutter and repetition, yielding a \emph{Regression Prism}.
%%
This opens up space for displaying additional information.
%%
Along the diagonal, $X$ and $Y$ assume the same feature, $Z \sim X + Y$ turns into $Z \sim X$ because the regression automatically ignores doublings.
%%
The diagonal therefore acts as reference on how strong the correlation for the given row (or column) feature is.

>>>>

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item The analysis of confounding features by providing comparison views between different formula results.
%	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
	\item The expansion and adaption of our method to any data with our open web-based system.
\end{itemize}
%%
The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{itemize}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{itemize}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow. Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and impose the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value is in the range $[0, 1]$, where 1 encodes a perfect fit.

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
\begin{equation}
Dependent \sim Independent_1 + ... + Independent_n
%Dependent \sim \sum_{i=1}^n Independent_i
\label{eq:RegressionNotation}
\end{equation}
%%
An example of a regression formula would be $KidneyDisorder \sim Smoking + Obesity$.
The most commonly used regression operators comprise:
%%
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable (e.g. $x\pm y$),
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
The class of the target feature restricts the regression type.
%%
Different regression types are available, we focus on the following:
%%
\paragraph{Linear Regression for Continuous Target.} The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
%%
The \emph{dependent} variable has to be of a continuous type.
%%

@@

While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
\begin{enumerate}
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
%%
The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
These target features often indicate the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing these relationships.
%%
% unified -> singular [Steffi?]
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
\subsection{Cube Description Using Regression Formula Notation}
%%
Expert knowledge is integrated into regression analysis by the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to
%%

@@

\subsection{Abstracting Regression Results Using $R^2$}
\begin{figure}[htb]
 \centering
 % \includegraphics[width=3.5in]{figures/cube}
 \includegraphics[width=1.0\linewidth]{figures/cube}
 \caption{
 (a) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The $R^2$ values extracted from the regression formulas depict the quality-of-fit and are mapped to color saturation (a saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The $R^2$ value of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%

\paragraph{3D (Cube) View.}
%%
Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube}b).
%%
$R^2$ values of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
The visualization of $R^2$ values derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
% e.g. wenn man es wâ°hlen kann
% i.e. wenn es so ist
Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., blue for linear regression and red for logistic regression).
%%
% Adverb immer vors Hauptverb oder Satzende
Thus, the cube can be easily extended using other regression types.
%%
For cubes with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
%The color for both 2D and 3D view encodes the type of the regression analysis (blue for logistic regressions, orange for linear regressions).
%%
\\\\
%%
Our goal is to create an overview visualization for a data set, but we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{Regression Cube}.

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{Regression Cubes} (based on~\cite{Keim}).
 %%
 % The analysis starts with a set of features given by the data set.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare regression cubes.
 %%
 [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
%\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
%\com{Workflow for different analysis approaches.}
%\com{0. Select the Variables.}
%\com{1. Select the Formulas}
%\com{a) Analysis without hypothesis about the data set (z ~ x+y)}
%\com{b) Analysis without hypothesis with confounder (z ~ age)}
%\com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
%\com{d) Use Difference View}
%\com{d) Clarify using confounder (Var ~ x)}
%\com{3. Find interesting pane.}
%\com{4. Get Details on Demand.}
%%
% â¹bergang ist recht holprig gewesen
Our \emph{Regression Cube} model is well-suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:

@@

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new regression cube for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
Input formulas reflect \emph{hypotheses} about the data.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

%%
$Cancer \sim X + Y + Z$ is the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
%%
All combinations of three independent features towards the target are analyzed through the cube.

@@

\section{System Design} \label{sec:SystemDesign}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/system}
 \caption{
 %%
 Breast fat data set loaded into the \emph{Regression Cube} prototype.
 %%
 (a) Using the formula input, the user specifies the dependent feature as well as its calculation rules.
 %%
 (b) 3D cube visualization, showing values above the cube matrix diagonal as overview.
 %%
 The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
 %%
 (c) 2D mosaic plot visualization of the selected slice for feature \emph{Pain/Discomfort}.
 %%††
 }
  \label{fig:System}
\end{figure}
%% scientific english: kurz und langweilig ;)
We designed our system to be freely accessible and easy to use.
With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
%%
The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.
%%
Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
%%
\\\\
%%
\noindent Design choices and spaces are restricted by the underlying technologies and devices.
%%
The exchange of data and methods using web-based technologies offers many advantages.
%%
There is no set-up time involved and domain experts can use the methods from any computer connected with the web.
%%
By design, web technology is based on a client-server architecture, making it easy to outsource computationally heavy tasks on server clusters and transferring results to the client device.
%%
The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
%%
Therefore, we have to adapt our design respecting the conventions of web pages.
%%
\subsection{System Paradigm and Components}
%%
The \emph{Regression Cube} design focuses on a clean interface, reducing the amount of user-interface elements as much as possible.
%%
This allows for a fast learning of the individual system parts.
%%
%More importantly, it allows the user to focus all mental resources on the analysis tasks, rather than wrangling with configuring the system to serve the current task or hypothesis.
%%
Our prototype consists of three components:
\begin{itemize}
	%%
	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
	%%
	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube [2.SI].
	%%
	\item The \emph{formula editor} allows formula input w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
	%%
	It also allows to select a reference formula for creating a difference cube [1.AF, 3.ZF].
	%%
\end{itemize}
%%\com{Focus on simplification and ease of use, reducing the number of UI elements as much as possible (Statistical Visualizations often tend to put all information and settings at once to the user, which is overwhelming for the user).}
\paragraph{File Upload and Classification [1.AF].}
%%
Popular analytics tools, such as WEKA \cite{WEKA}, owe their popularity to their support of open file types.
%%
To allow other users even outside the epidemiological application domain to access to our tool, we use standard ASCII-based CSV files.
%%
The first line in a CSV file represents all features (columns) of the data set.
%%
Each line after that represents one subject (row) and its feature manifestations.

\textbf{Encoding via csv Files.}
%%
Encoding variable types in CSV files is not standardized.
%%
However, we need to ensure the correct variable type classification and have to enforce some basic standards.
%%
All categorical values have to be enclosed by quotation marks.
%%
Continuous variables are denoted as digits without enclosing quotation marks.
%%
Although this seems obvious, many cohort study data sets encode categorical features using ID values that are denoted in a data dictionary.
%%
Variables with only two manifestations are classified as dichotomous, leading to three possible data types: numerical, categorical and categorical/dichotomous.
%%
Missing values are denoted by using no character at all, a whitespace, or an empty quotation mark encapsulated string.

@@

\paragraph{Formula Editor [1.AF, 3.ZF].}
%%
After uploading the data, the user can specify a formula or use the default ($Z \sim X + Y$).
%%
Entering a formula is facilitated via text input.
%%
On formula input, a context modal displays all data set features as well as the available operators and their function.
%%
This allows to comprehend the function of the underlying formula for users without statistical background about regression analysis and its notation.
%%
Auto-completing input features also simplifies the approach and works as spell check of feature names.
%%

@@

\paragraph{Usage of a Regression Prism for Information Reduction.}
%%
Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the mosaic plot matrix.
%%
This is due to the symmetry of basic regression operators.
%%
$Z \sim X + Y$ produces the same result as $Z \sim Y + X$.
%%
Therefore, we can discard half of the results to reduce visual clutter and repetition, yielding a \emph{Regression Prism}.
%%
This opens up space for displaying additional information.
%%
Along the diagonal, $X$ and $Y$ assume the same feature, $Z \sim X + Y$ turns into $Z \sim X$ because the regression automatically ignores doublings.
%%
The diagonal therefore acts as reference on how strong the correlation for the given row (or column) feature is.

\paragraph{3D Prism as Data Mini-Map.}
%%
The 3D cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
%%
It serves as a function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
%%
%It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
%%
The displayed prism shows values above the matrix diagonal.
%%
For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System}b).
%%
Applying this strategy to a formula containing a static target (e.g. $Cancer \sim X + Y + Z$) yields many occlusions, since the CFS algorithm creates the same feature space for every slice.
%%
For such formulas, the 3D view encodes every data element as absolute difference between its $R^2$ features and the global mean along the z-axis.
%%
This highlights slices with unusually low or high results (Fig.~\ref{fig:Application}).
%%
\paragraph{Tackling the disadvantages of 3D information visualization.}
%%
3D information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
%%
We aim to minimize these problems as much as possible.
%%
Since the $R^2$ values are mapped on data point opacity, large values are highlighted in the prism, guiding the focus to the respective slices.
%%
It also creates a sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
%%
Also, the preceding correlation-based feature selection reduces the information space significantly, leading to sparse cubes.
%%
Overlapping is still an issue, but this way greatly reduced in its affect to the readability of the visualization.
%%
